%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Graduate Curriculum Vitae
% LaTeX Template
% Version 1.1 (9/12/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
% Important note:
% This template requires the res.cls file to be in the same directory as the
% .tex file. The res.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[margin, 10pt]{res} % Use the res.cls style, the font size can be changed to 11pt or 12pt here

\usepackage{helvet} % Default font is the helvetica postscript font
%\usepackage{newcent} % To change the default font to the new century schoolbook postscript font uncomment this line and comment the one above

\setlength{\textwidth}{5.1in} % Text width of the document

\newsectionwidth{1.5in} % How much horizontal space titles have

\headheight = -10pt % Move up top of page 10pts

\pagestyle{plain} % Page style with page numbers at center-bottom of every page

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }

\begin{document}

%----------------------------------------------------------------------------------------
%	NAME AND CONTACT INFORMATION SECTION
%----------------------------------------------------------------------------------------

\moveleft.5\hoffset\centerline{\Large\bf Nathan Riley Summers LaPierre} % Your name at the top
\moveleft.5\hoffset\centerline{NathanRLaPierre.com}
\moveleft.5\hoffset\centerline{NathanL2012@gmail.com}
%\moveleft.5\hoffset\centerline{(571) 839-5008}
 
\moveleft\hoffset\vbox{\hrule width\resumewidth height 1pt}\smallskip % Horizontal line after name; adjust line thickness by changing the '1pt'

%----------------------------------------------------------------------------------------

\begin{resume}

%----------------------------------------------------------------------------------------
%	POSITIONS & EMPLOYMENT SECTION
%----------------------------------------------------------------------------------------

\section{CURRENT POSITION}

{\sl Postdoctoral Research Fellow} \hfill Mar. 2025 - Present \\ Department of Medical Oncology \\ Dana-Farber Cancer Institute \\ Lab of Alexander Gusev \\


\section{OTHER RECENT POSITIONS}

{\sl Postdoctoral Scholar and Fellow} \hfill Aug. 2022 - Mar. 2025 \\ Department of Human Genetics \\ University of Chicago \\ Labs of Xin He and Matthew Stephens



%----------------------------------------------------------------------------------------
%	EDUCATION SECTION
%----------------------------------------------------------------------------------------

\section{EDUCATION}

PhD in Computer Science, June 2022 \\
University of California, Los Angeles \\
Labs of Eleazar Eskin and Wei Wang

M.S. in Computer Science, March 2019 -- 3.74 GPA \\
University of California, Los Angeles

B.S. in Applied Computer Science, December 2015 -- 3.86 GPA ({\sl magna cum laude}) \\
George Mason University

%----------------------------------------------------------------------------------------
%	RECENT COMPETITION RESULTS SECTION
%----------------------------------------------------------------------------------------

\section{RECENT COMPETITION RESULTS}

Arc Institute Virtual Cell Challenge (Nov. 2025) \\
1st in Pearson Correlation, 10th in Generalist Rankings, 38th Overall (out of 1265+ total entrants in validation and final rounds)
  
%----------------------------------------------------------------------------------------
%	PUBLICATIONS SECTION
%----------------------------------------------------------------------------------------

\section{PUBLICATIONS}

\bibliography{references} 
\bibliographystyle{ieeetr}

%Mohammad Arifur Rahman, \textbf{Nathan LaPierre}, Huzefa Rangwala, and Daniel Barbara, ``Clustering Metagenome Sequences Using Canopies", in {\sl 9th International Conference on Bioinformatics and Computational Biology}, Honolulu, Hawaii, 2017.

For all research software, please see: \href{https://github.com/nlapier2/}{https://github.com/nlapier2/} \\
Google Scholar: \href{https://scholar.google.com/citations?user=EVMceKUAAAAJ\&hl=en}{https://scholar.google.com/citations?user=EVMceKUAAAAJ\&hl=en}


Z. Lu, Y. Ding, \textbf{N. LaPierre}, et al, ``Integrating perturbational screens, eQTL, and GWAS data identifies mediating genes for complex traits,'' \emph{medRxiv}, Jan. 2026.

A. Khan, J. Bagley, \textbf{N. LaPierre}, et al, ``Differing Genetics of Saline and Cocaine Self‐Administration in the Hybrid Mouse Diversity Panel,'' \emph{Genes, Brain and Behavior}, vol. 24, no. 3, pp. e70029, June 2025.

P. B. Chen, R. Chen, \textbf{N. LaPierre}, et al, ``Complementation testing identifies genes mediating effects at quantitative trait loci underlying fear-related behavior,'' \emph{Cell Genomics}, vol. 4, no. 5, pp. 100545, May 2024. PubMed PMID: 38697120.

M. Alser, ..., \textbf{N. LaPierre}, ..., ``Packaging and containerization of computational methods,'' \emph{Nature Protocols}, vo. 19, no. 9, pp. 2529–-2539, Apr. 2024. PubMed PMID: 38565959.

\textbf{N. LaPierre} and Harold Pimentel, ``Accounting for Isoform Expression in eQTL Mapping Substantially Increases Power,'' \emph{PLOS Computational Biology}, vol. 20, no. 2, pp. e1011857, Feb. 2024. PubMed PMID: 38346082. PubMed Central PMCID: PMC10890775.

A. Khan, J. Bagley, \textbf{N. LaPierre}, et al, ``Genetic pathways regulating the longitudinal acquisition of cocaine self-administration in inbred and recombinant inbred mice,'' \emph{Cell Reports}, vol. 42, no. 8, pp. 112856, Aug. 2023.  PubMed PMID: 37481717; PubMed Central PMCID: PMC10530068.  

\textbf{N. LaPierre}*, B. Fu*, S. Turnbull, E. Eskin, and S. Sankararaman, ``Leveraging family data to design Mendelian Randomization that is provably robust to population stratification,'' \emph{Genome Research}, vol. 33, pp. 1032-1041, July 2023.  PubMed PMID: 37197991; PubMed Central PMCID: PMC10538495.

B. Fuqua, ... \textbf{N. LaPierre}, ..., ``The Genetic Architecture of Dietary Iron Overload and Associated Pathology in Mice,'' \emph{bioRxiv}, June 2023.

J. Karlin, L. Gai, \textbf{N. LaPierre}, et al, ``Ensemble neural network model for detecting thyroid eye disease using external photographs,'' \emph{British Journal of Ophthalmology}, Sep. 2022.  PubMed PMID: 36126104.

F. Meyer, ... \textbf{N. LaPierre}, ..., ``Critical Assessment of Metagenome Interpretation-the second round of challenges,'' \emph{Nature Methods}, vol. 19, pp. 429–440, Apr. 2022.  PubMed PMID: 35396482; PubMed Central PMCID: PMC9007738.

C. Cinelli, \textbf{N. LaPierre}, B. L. Hill, S. Sankararaman, and E Eskin, ''Robust Mendelian randomization in the presence of residual population stratification, batch effects and horizontal pleiotropy,'' \emph{Nature Communications}, vol. 13, no 1., pp. 1093, Mar. 2022.  PubMed PMID: 35232963; PubMed Central PMCID: PMC8888767.

\textbf{N. LaPierre}*, K. Taraszka*, H. Huang, R. He, F. Hormozdiari, and E. Eskin, ``Identifying Causal Variants by Fine Mapping Across Multiple Studies,'' \emph{PLOS Genetics}, vol. 17, no. 9, pp. e1009733, Sept. 2021.  PubMed PMID: 34543273; PubMed Central PMCID: PMC8491908.

J. Bloom et al (many authors), ``Massively scaled-up testing for SARS-CoV-2 RNA via next-generation sequencing of pooled and barcoded nasal and saliva samples,'' \emph{Nature Biomedical Engineering}, vol. 5, no. 7, pp. 657-665, July 2021.  PubMed PMID: 34211145.

\textbf{N. LaPierre}, M. Alser, E. Eskin, D. Koslicki*, and S. Mangul*, ``Metalign: Efficient alignment-based metagenomic profiling via containment min hash,'' \emph{Genome Biology}, vol. 21, pp. e242, Sep. 2020. PubMed PMID: 32912225; PubMed Central PMCID: PMC7488264.

%\textbf{N. LaPierre}*, K. Collins*, J. Rotman, and E. Eskin, ``Identifying Causal Variants by Fine Mapping Across Multiple Studies,'' in {\sl International Conference on Research in Computational Molecular Biology (RECOMB)}, virtual, June 2020, pp. 257-258.

\textbf{N. LaPierre}, R. Egan, W. Wang, and Z. Wang, ``De novo Nanopore read quality improvement using deep learning,'' \emph{BMC Bioinformatics}, vol. 20, no. 1, pp. e552, Dec. 2019. PubMed PMID: 31694525; PubMed Central PMCID: PMC6833143.

\textbf{N. LaPierre}, C. Ju, G. Zhou, and W. Wang, ``MetaPheno: A Critical Evaluation of Deep Learning and Machine Learning in Metagenome-Based Disease Prediction,'' \emph{Methods}, vol. 166, pp. 74-82, Aug. 2019. PubMed PMID: 30885720; PubMed Central PMCID: PMC6708502.

\textbf{N. LaPierre}*, S. Mangul*, M. Alser, I. Mandric, N.C. Wu, D. Koslicki, and E. Eskin, ``MiCoP: Microbial Community Profiling method capable of detecting low abundance viral and fungal organisms in metagenomic samples,'' \emph{BMC Genomics}, vol. 20, no. 5, pp. e423, June 2019. PubMed PMID: 31167634; PubMed Central PMCID: PMC6551237.

M. A. Rahman, \textbf{N. LaPierre}, H. Rangwala, and D. Barbara, ``Metagenome sequence clustering with hash-based canopies,'' \emph{Journal of bioinformatics and computational biology}, vol. 15, no. 6, pp.1740006, Oct. 2017. PubMed PMID: 29113561.

M. A. Rahman, \textbf{N. LaPierre}, and H. Rangwala, ``Phenotype Prediction from Metagenomic Data Using Clustering and Assembly with Multiple Instance Learning (CAMIL),'' \emph{IEEE/ACM transactions on computational biology and bioinformatics}, Oct. 2017. PubMed PMID: 28981422.

\textbf{N. LaPierre}, M. A. Rahman, and H. Rangwala, ``CAMIL: Clustering and Assembly with Multiple Instance Learning for Phenotype Prediction,'' in {\sl IEEE International Conference on Bioinformatics and Biomedicine}, Shenzhen, China, 2016.

\textbf{N. LaPierre} and H. Rangwala, ``Predicting Clinical Phenotype using OTU-based Metagenome Representation,'' in {\sl IEEE International Conference on Data Mining Workshop on Biological Data Mining and its Applications in Healthcare}, Atlantic City, New Jersey, 2015, pp. 156-163.

(*  Authors contributed equally)



%----------------------------------------------------------------------------------------
%	RESEARCH PRESENTATIONS SECTION
%----------------------------------------------------------------------------------------

\section{CONFERENCE TALKS}

``Pleiotropy-robust methods for high-dimensional multivariable Mendelian randomization,'' in {\sl STATGEN 2024: Conference on Statistics in Genomics and Genetics}, May 2024.

``Leveraging family data to design Mendelian Randomization that is provably robust to population stratification,'' in {\sl International Conference on Research in Computational Molecular Biology (RECOMB)}, Apr. 2023.

``Metalign: Efficient alignment-based metagenomic profiling via containment min hash,'' in {\sl Intelligent Systems for Molecular Biology (ISMB) HitSeq}, virtual, July 2020.

``Identifying Causal Variants by Fine Mapping Across Multiple Studies'' in {\sl International Conference on Research in Computational Molecular Biology (RECOMB)}, virtual, June 2020.

``Metalign: Efficient alignment-based metagenomic profiling via containment min hash,'' in {\sl RECOMB-Seq}, virtual, June 2020.

``CAMIL: Clustering and Assembly with Multiple Instance Learning for Phenotype Prediction'' in IEEE International Conference on Bioinformatics and Biomedicine, December 2016.

``Predicting Clinical Phenotype using OTU-based Metagenome Representation'' in IEEE International Conference on Data Mining workshop on Biological Data Mining and its Applications in Healthcare, November 2015.


\section{CONFERENCE POSTERS}

``Accounting for Isoform Expression in eQTL Mapping,'' in {\sl CSHL Genome Informatics}, virtual, Nov. 2021.

``Identifying Causal Variants by Fine Mapping Across Multiple Studies'' in American Society for Human Genetics Annual Meeting, October 2019. 

``MiniScrub: de novo long read scrubbing using approximate alignment and deep learning'' in Amazon Web Services - UCLA Computational Medicine Symposium, February 2019.

``MiniScrub: de novo long read scrubbing using approximate alignment and deep learning'' in American Society for Human Genetics Annual Meeting, October 2018.

``MiCoP: Microbial Community Profiling method capable of detecting low abundance viral and fungal organisms in metagenomic samples'' in American Society for Human Genetics Annual Meeting, October 2017.

``Developing a Computational Pipeline for Metagenomic State Classification with Feature Engineering'' in Volgenau School of Engineering Undergraduate Research Celebration, April 2015
\begin{itemize} \itemsep -2pt % Reduce space between items
%\item Presented an earlier iteration of my research project in poster form
\item Won Outstanding Undergraduate Research Project Award for poster
\end{itemize}



%----------------------------------------------------------------------------------------
%	AWARDS and HONORS SECTION
%----------------------------------------------------------------------------------------

\section{FUNDING AND AWARDS}

Fellowships and Merit-Based Scholarships
\begin{itemize} \itemsep -2pt % Reduce space between items
\item {\sl NIH T32 HL007605 - Research Training in Respiratory Biology}, July 2023-2024, University of Chicago
\item {\sl ModEling and uNdersTanding human behaviOR (MENTOR) NSF Training Grant}, 2018-19 academic year at UCLA
\item {\sl NIH T32 Doctoral Training Fellowship}, 2016-18 academic years at UCLA
\item {\sl Bersoff Endowed Scholarship}, Outstanding Academic Achievement, Awarded twice in 2015 and 2014
\item {\sl SWIFT Scholarship}, Outstanding Academic Achievement, Awarded in 2015
\end{itemize}

Selected Honors
\begin{itemize} \itemsep -2pt % Reduce space between items
\item {\sl Honorable Mention in NSF Graduate Research Fellowship}, 2015-16
\item {\sl Outstanding Undergraduate Student Award}, given to overall best undergraduate student in Computer Science at George Mason University, May 2016
\item {\sl Outstanding Academic Achievement Award}, given for outstanding performance in Computer Science, May 2016
\item {\sl Outstanding Undergraduate Research Project} for presentation of my research project at Volgenau School of Engineering Undergraduate Research Symposium, April 2015
\item {\sl Outstanding Undergraduate Teaching Assistant} for two semesters of excellence as a teaching assistant, April 2015
\item {\sl Dean's List} every semester
\end{itemize}


Technical Competitions
\begin{itemize} \itemsep -2pt % Reduce space between items
\item {\sl Top 5 at VTHacks}, a software development competition at Virginia Tech with over 45 teams, April 2014
\item {\sl 2nd Place in the Technical Innovation Challenge}, a week-long competition at George Mason University to design a viable software product, jointly refereed by Computer Science and Business Departments, October 2014
\end{itemize}



%----------------------------------------------------------------------------------------
%	TEACHING EXPERIENCE SECTION
%----------------------------------------------------------------------------------------
 
\section{TEACHING EXPERIENCE}

{\sl Graduate Teaching Assistant} \hfill Spring 2020 \\
CS/BIOINFO 122/222 - Algorithms in Bioinformatics, Computer Science Department, UCLA

{\sl Graduate Teaching Assistant} \hfill Winter 2020 \\
CS/BIOINFO 124/224 - Machine Learning in Genomics, Computer Science Department, UCLA

{\sl Undergraduate Teaching Assistant} \hfill Fall 2014 - Spring 2015 \\
CS 306 - Computer Law and Ethics, Computer Science Department, George Mason University
\begin{itemize} \itemsep -2pt % Reduce space between items
\item Received Outstanding Undergraduate Teaching Assistant Award 
\item Assisted students with legal research, writing, and oral communication
\item Responsible for grading student assignments worth 25\% of their grade; one of the only Undergraduate Teaching Assistants entrusted with this responsibility
\end{itemize} 



%----------------------------------------------------------------------------------------
%	VOLUNTEERING AND LEADERSHIP SECTION
%----------------------------------------------------------------------------------------

\section{VOLUNTEERING, ORGANIZATIONS, AND LEADERSHIP}
{\sl Diversity, Equity, and Inclusion Committee -- University of Chicago Departments of Human Genetics and Genetics, Genomics, \& Systems Biology} \hfill 2023 - present
\begin{itemize} \itemsep -2pt % Reduce space between items
\item Participated in monthly discussions to plan and organize events and resources promoting DEI initiatives for the HG and GGSB departments
\item Currently revising a yearly survey to be sent out to survey the DEI ``climate'' in these departments
\end{itemize}

{\sl Student-Run Computing and Technology (SRCT)} \hfill 2013 - 2016 \begin{itemize} \itemsep -2pt % Reduce space between items
\item Student organization at George Mason University that works on software projects and competitions to benefit the university and broader local community
\item Secretary and Member of Executive Board, Fall 2015 Semester
\end{itemize}

Executive Curriculum Planner and Mentor, Community Programming Initiative \\ George Mason University Honors College and SRCT
 \hfill 2014 - 2015
\begin{itemize} \itemsep -2pt % Reduce space between items
\item Volunteer effort to teach basic programming to local 5th-8th grade students
\item Planned and developed parts of the curriculum for the sessions, such as designing games and hints to help the students create them
\item Mentored the elementary and middle school students during the sessions
\end{itemize} 



%----------------------------------------------------------------------------------------
%	COMPUTATIONAL SKILLS SECTION
%----------------------------------------------------------------------------------------

\section{COMPUTATIONAL \\ SKILLS} 

{\sl Languages / Scripting:} 
Fluent in Python, R, Bash; past experience with C(++) and Java\\
{\sl Deep Learning Frameworks:} PyTorch, Keras \\
{\sl Other Technologies:} Git, Docker, LaTeX \\
{\sl Bioinformatics experience:} Single cell RNA-seq and Perturb-seq; Mendelian randomization; GWAS \& statistical fine mapping; alignment; assembly; metagenomics; common software such as Scanpy, AnnData, Plink, Bedtools, Samtools, Bcftools, etc; SGE and Slurm grid/cluster computing systems \\
{\sl Machine Learning Algorithms:} Deep Learning, Linear \& Logistic Regression, SVMs, Random Forests, etc \\
{\sl Operating Systems:} Linux, Windows, macOS 



%----------------------------------------------------------------------------------------
%	PROFESSIONAL EXPERIENCE SECTION
%----------------------------------------------------------------------------------------
 
\section{INTERNSHIPS}

{\sl Graduate Student Researcher} \hfill Summer 2017 \\ DOE Joint Genome Institute / Lawrence Berkelely National Lab

\begin{itemize} \itemsep -2pt % Reduce space between items
\item Developed a method that uses a Convolutional Neural Network (Deep learning method) to improve Oxford Nanopore long read quality \emph{de novo}
\item Method improves both read accuracy and downstream \emph{de novo} assembly
\item Helped pioneer use of deep learning within the lab
\end{itemize}


{\sl Security Engineering Intern} \hfill Summers 2013-2015 \\
Sony Corporation of America 

\begin{itemize} \itemsep -2pt % Reduce space between items
\item Wrote secure and scalable software and worked with Big Data in order to help analyze, detect, and prevent attacks on Sony\textsc{\char13}s networks
\item Used Python, Javascript (Node, Express, Meteor), MongoDB, X/HTML, CSS 
\item Developed a workplace communications system using Meteor.js; solo project
\item Assisted in the development of a web application based on a searchable database system using Node.js, Express.js, and MongoDB
\item Assisted in the development of a network forensics system; wrote backend python scripts and XML web layouts
\end{itemize} 



%----------------------------------------------------------------------------------------

\end{resume}
\end{document}